Cognate Priming: Report

### 2025-12-09 docs/manuscript.qmd
Fix code

targets::tar_load(c(participants,
					stimuli,
					vocabulary,
					attrition_trials,
					attrition_participants,
					bvq_data,
					gaze,
					data_bcn,
					data_oxf))

targets::tar_load(c(model_fits_bcn,
					model_loos_bcn,
					model_fits_oxf,
					model_loos_oxf))

# prepare data
attrition <- lst(participants,
				 attrition_participants,
				 attrition_trials) |>
	reduce(inner_join, by = join_by(session_id)) |>
	unnest_wider(matches("all"), names_sep = "_") |>
	rename_with(\(x) gsub("is_valid_gaze_all_|is_valid_vocab_all",
						  "", x))

@@

options(ggplot2.ordinal.fill = clrs,
		ggplot2.ordinal.colour = clrs,
		ggplot2.discrete.fill = clrs,
		ggplot2.discrete.colour = clrs,
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

@@

ages <- participants_oxf |>
	summarise(n = n(),
			  n_participants = n_distinct(child_id),
			  across(age, lst(mean, sd, min, max)),
			  across(matches("doe_"), lst(mean, sd)),
			  age_vctr = list(age),
			  .by = c(lp, test_language))

suppressWarnings({
	time_diffs <- participants_oxf |>
		arrange(child_id, session_id, age) |>
		mutate(n_session = 1:n()) |>
		select(child_id, session_id, session_n, age) |>
		filter(child_id %in% participants_oxf$child_id[duplicated(participants_oxf$child_id)]) |>
		pivot_wider(id_cols = child_id,
					names_from = session_n,
					values_from = age,
					names_repair = janitor::make_clean_names) |>
		rowwise() |>
		mutate(diff_min = x2 - x1) |>
		ungroup() |>
		summarise(diff_min = min(diff_min, na.rm = TRUE))
})

@@

```{r tbl-participants}
#| label: tbl-participants
#| tbl-cap: "Demographic and linguistic profile of testing sessions. The number of excluded testing sessions and participants is indicated between parentheses."
participants |>
	unnest_wider(doe) |>
	select(-matches("others")) |>
	arrange(lp, test_language) |>
	distinct(session_id, child_id, .keep_all = TRUE) |>
	mutate(study = if_else(location=="Oxford", 
						   "Study 1",
						   "Study 2"),
		   test_language = paste0(test_language, " dominant"),
		   lp = gsub(" \\(English\\)", "", lp)) |> 
	inner_join(attrition_participants,
			   by = join_by(session_id)) |> 
	summarise(n = n(),
			  n_participants = length(unique(child_id)),
			  across(age, lst(mean, sd)),
			  across(matches("doe_"), lst(mean, sd)),
			  # age_vctr = list(age),
			  .by = c(lp, study, test_language, is_valid_participant)) |> 
	pivot_wider(id_cols = c(lp, study, test_language),
				names_from = is_valid_participant,
				values_from = is.numeric) |> 
	select(lp, study, test_language, starts_with("n_"), matches("TRUE"),
		   n_FALSE, n_participants_FALSE) |> 
	rename_with(\(x) gsub("_TRUE", "", x)) |> 
	gt(rowname_col = "test_language",
	   groupname_col = c("study", "lp")) |>
	grand_summary_rows(columns = matches("age"),
					   fns = list(Total ~ mean(.)),
					   fmt = ~ fmt_number(., decimals = 2)) |>
	grand_summary_rows(columns = c(n, n_participants),
					   fns = list(Total ~ sum(.)),
					   fmt = ~ fmt_integer(.)) |>
	# gt_plt_dist(age_vctr, type = "histogram") |>
	fmt_number(is.numeric, decimals = 2) |>
	fmt_integer(c(n, n_participants, n_FALSE, n_participants_FALSE)) |>
	fmt_number(matches("doe"), decimals = 2, scale_by = 100) |>
	cols_merge(c(n, n_FALSE), pattern = "{1} ({2})") |>
	cols_merge(c(n_participants, n_participants_FALSE),
			   pattern = "{1} ({2})") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Test sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

```{r vocab-values-oxf}
vocabulary_oxf <- inner_join(vocabulary,
							 select(participants_oxf, session_id),
							 by = join_by(session_id))

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. For visualisation purposes, testing sessions were aggregated at three age points: 21, 25, and 30 months. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 5
#| fig-height: 4
#| out-width: 70%
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, age,
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	filter(is_valid_participant) |> 
	select(child_id, session_id, vocab_id, is_imputed, l1_count,
		   lp, location, age_group, age) |> 
	mutate(location = factor(location, 
							 levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

vocab_plot_means <- vocab_plot_data |> 
	summarise(l1_count_sd = sd(l1_count, na.rm = TRUE),
			  l1_count = mean(l1_count, na.rm = TRUE),
			  n = n(),
			  .by = c(age, side, lp, study)) |> 
	mutate(.lower = l1_count - (l1_count_sd/sqrt(n)),
		   .upper = l1_count + (l1_count_sd/sqrt(n)))

vocab_plot_data |> 
	filter(study=="Study 1 (English)") |> 
	ggplot(aes(age, l1_count, 
			   side = side)) +
	geom_point(size = 2,
			   shape = 1,
			   colour = "grey",
			   alpha = 3/4,
			   stroke = 1) +
	geom_smooth(method = "lm", 
				formula = "y ~ x",
				colour = "black",
				fill = "black") +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(limits = c(0, 575),
					   breaks = seq(0, 1e3, 100)) +
	scale_x_continuous(breaks = seq(0, 50, 2)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```

@@

Participants were presented with 32 trials in random order, which belonged to three conditions: *Unrelated* (16 trials), *Related/Non-cognate* (8 trials), and *Related/Cognate* (8 trials). In *Unrelated* trials, target was phonologically unrelated with the English and Spanish labels of the prime picture (e.g., $\text{door}_{\text{ENG}}/\text{puerta}_{\text{SPA}}-\text{sock}_{\text{ENG}}$). In *Related/Non-cognate* trials, the target shared phonological onset with the English prime label, but not with the Spanish prime label (e.g., $\textbf{t}\text{ree}_{\text{ENG}}/\text{árbol}_{\text{SPA}}-\textbf{t}\text{ruck}_{\text{ENG}}$). In *Related/Cognate* trials, the target shared phonological overlap with both English and Spanish prime labels ($\textbf{b}\text{oot}_{\text{ENG}}/\textbf{b}\text{ota}_{\text{SPA}}-\textbf{b}\text{ird}_{\text{ENG}}$). Especial attention was paid to avoiding semantic or taxonomic relationships between prime and target words, and between prime and distractor words. Distractors were always phonologically unrelated to the prime and target labels in the same trial.

@@

```{r stim-stats}
durations <- stimuli |>
	summarise(across(duration, 
					 lst(mean, sd, min, max)), 
			  .by = c(test_language)) %>%
	mutate(across(is.numeric, \(x) format(round(x*1e3, 2), big.mark = ","))) %>%
	split(janitor::make_clean_names(.$test_language))
```

@@

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type. Values are summarised using the mean and the standard deviation (between parentheses)."
stimuli |>
	unnest_wider(c(freq, nphon, familiarity)) |>
	summarise(across(c(matches("n_phon|freq_|nphon|familiarity")),
					 list(	mean = \(x) mean(x, na.rm = TRUE),
					 	  sd = \(x) sd(x, na.rm = TRUE))),
			  .by = c(trial_type, test_language)) |>
	select(-matches("distractor")) |>
	mutate(test_language = if_else(test_language=="English",
								   paste0("Study 1: ", test_language),
								   paste0("Study 2: ", test_language))) |>
	arrange(test_language) |> 
	gt(rowname_col = "trial_type",
	   groupname_col = "test_language") |> 
	# summary_rows(groups = c("Unrelated", "Non-cognate", "Cognate"),
	# 			 columns = is.numeric,
	# 			 fns = list(Total ~ mean(.)),
	# 			 fmt = ~ fmt_number(., decimals = 2)) |>
	grand_summary_rows(columns = matches("phon|freq"),
					   fns = list(Total ~ mean(.)),
					   fmt = ~ fmt_number(., decimals = 2)) |>
		grand_summary_rows(columns = matches("familiarity"),
					   fns = list(Total ~ mean(.)),
					   fmt = ~ fmt_number(., decimals = 2, scale_by = 100)) |>
	tab_spanner("# Phonemes", matches("nphon")) |>
	tab_spanner("Frequency (Zipf)", matches("freq")) |>
	tab_spanner("Familiarity (%)", matches("familiarity")) |>
	fmt_number(is.numeric, decimals = 2) |>
	fmt_number(matches("familiarity"), decimals = 2, scale_by = 100) |> 
	cols_merge(c(freq_prime_mean, freq_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(freq_target_mean, freq_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(familiarity_prime_mean, familiarity_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(familiarity_target_mean, familiarity_target_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_prime_mean, nphon_prime_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(nphon_target_mean, nphon_target_sd),
			   pattern = "{1} ({2})") |>
	cols_label(trial_type = "",
			   familiarity_prime_mean = "Prime",
			   familiarity_target_mean = "Target",
			   freq_prime_mean = "Prime",
			   freq_target_mean = "Target",
			   nphon_prime_mean = "Prime",
			   nphon_target_mean = "Target") |>
	tab_style(list(cell_text(style = "italic")), 
			  list(cells_column_labels(),
			  	 cells_column_spanners(),
			  	 cells_grand_summary()))
```

@@

We defined a time window of interest from 200 ms after target and distractor pictures onset until the end of the trial at 2,000 ms when both pictures disappeared from screen. The first 200 ms of the test phase were discarded to avoid modelling fixations driven by processes other than auditory word recognition [@fernald1998rapid; @fernald2001half]. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of 166.67 ms). Target looking probability was calculated as the empirical logit, using the number of samples inside the time bin in which the participant was looking at the target and distractor AOI (see @eq-elogit) [@barr2008analyzing; @agresti2012categorical], as follows:

@@

After trials that matched any of the aforementioned exclusion criteria from the data set, we excluded participants who did not provide at least two valid trials in each condition (*n* = `r n_exc_participants`), and participants with a vocabulary size lower than 42, which corresponds to 10% of the words in the OCDI vocabulary checklist (*n* = `r n_exc_participants_vocab`). The final data set included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final data set, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

@@

n_lp <- attrition |>
	distinct(is_valid_participant, lp, session_id) |>
	count(is_valid_participant, lp) |>
	pivot_wider(names_from = is_valid_participant,
				values_from = n,
				names_repair = janitor::make_clean_names) |>
	mutate(n_participants_true = true,
		   n_participants_false = false) |>
	select(lp, true, false)

attrition |>
	mutate(study = if_else(location=="Oxford", 
						   "Study 1: English (Oxford)",
						   "Study 2: Catalan/Spanish (Barcelona)")) |> 	
	summarise(n_trials_true = sum(is_valid_trial),
			  n_trials_false = sum(!is_valid_trial),
			  .by = c(lp, trial_type, study)) |>
	arrange(study, desc(lp)) |>
	pivot_wider(id_cols = c(study, lp),
				names_from = c(trial_type),
				values_from = matches("trials"),
				names_repair = janitor::make_clean_names) |>
	left_join(n_lp, by = join_by(lp)) |>
	select(study, lp, true, false, matches("n_trials")) |>
	gt(rowname_col = "lp",
	   groupname_col = "study") |>
	tab_spanner("Participants", matches("participant")) |>
	tab_spanner("Trials", matches("trial")) |>
	fmt_integer(everything()) |>
	cols_merge(c(true, false), pattern = "{1} ({2})") |>
	cols_merge(c(n_trials_true_cognate, n_trials_false_cognate),
			   pattern = "{1} ({2})") |>
	cols_merge(c(n_trials_true_non_cognate, n_trials_false_non_cognate),
			   pattern = "{1} ({2})") |>
	cols_merge(c(n_trials_true_unrelated, n_trials_false_unrelated),
			   pattern = "{1} ({2})") |>
	cols_label(true = "Participants",
			   n_trials_true_cognate = "Cognate",
			   n_trials_true_non_cognate = "Non-cognate",
			   n_trials_true_unrelated = "Unrelated") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(align = "left"),
			  cells_stub()) |>
	grand_summary_rows(fns = list(label = md("*N*")) ~ sum(.),
					   fmt = ~ fmt_integer(.))
```

@@

r_version <- glue::glue(r_info$R.version$major,
						r_info$R.version$minor,
						.sep = ".")
stan_version <- "2.33.0"

@@

We used Bayesian Hierarchical General Additive Models (HGAMs) to model the data [@pedersen2019hierarchical], using a Gaussian distribution to model the the logit of target looking. First, we fit a model ($\mathcal{M}_0$) that included the main effects of *Condition* and *Age* as fixed effects in the model. Given that the cognate/non-cognate distinction is meaningless for monolinguals, for the following analyses we lumped the *Related/Cognate* and *Related/Non-cognate* into a single level, *Related*. We set an *a priori* contrast for the *Condition* predictor [@schad2020capitalize], comparing *Unrelated* and *Related* trials (sum-coded as `-0.5` and `+0.5`. Before entering the model, the *Age* predictor was standardised by subtracting from each observation the mean of the predictor, and dividing the result from the standard deviation of the predictor.

We included the variable *Session*---which indexes individual testing sessions that may belong to the same participant---as grouping variable, nested within the *Participant* grouping variable---which indexes a distinct participant. This nested random effects structure incorporates the longitudinal design of data collection, in which multiple participants were tested more than once at different ages. We added by-session intercepts and *Condition* slopes, and by-participant intercepts and *Age* slopes. To model the time course of target looking across time bins, we included B-splines for the main effect of *Time*, and one for the interaction between *Group* and *Relatedness* [@wood2017generalized]. For both splines, we specified $k = 8$ basis functions or *knots*. @eq-model shows a formal implementation of the model.

@@

```{r coefs-oxf}
var_labels <- c("b_Intercept" = "Intercept",
				"b_age_std" = "Age (+1 SD)",
				"b_condition1" = "Condition (U vs. R/NC)",
				"b_condition2" = "Condition (R/NC vs. R/C)")

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

>>>>

targets::tar_load(c(
    participants,
    stimuli,
    vocabulary,
    attrition_trials,
    attrition_participants,
    bvq_data,
    gaze,
    data_bcn,
    data_oxf
))

targets::tar_load(c(
    model_fits_bcn,
    model_loos_bcn,
    model_fits_oxf,
    model_loos_oxf
))

# prepare data
attrition <- lst(
    participants,
    attrition_participants,
    attrition_trials
) |>
    reduce(inner_join, by = join_by(session_id)) |>
    unnest_wider(matches("all"), names_sep = "_") |>
    rename_with(\(x) gsub(
        "is_valid_gaze_all_|is_valid_vocab_all",
        "", x
    ))

@@

options(
    ggplot2.ordinal.fill = clrs,
    ggplot2.ordinal.colour = clrs,
    ggplot2.discrete.fill = clrs,
    ggplot2.discrete.colour = clrs,
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)

@@

ages <- participants_oxf |>
    summarise(
        n = n(),
        n_participants = n_distinct(child_id),
        across(age, lst(mean, sd, min, max)),
        across(matches("doe_"), lst(mean, sd)),
        age_vctr = list(age),
        .by = c(lp, test_language)
    )

suppressWarnings({
    time_diffs <- participants_oxf |>
        arrange(child_id, session_id, age) |>
        mutate(n_session = 1:n()) |>
        select(child_id, session_id, session_n, age) |>
        filter(child_id %in% participants_oxf$child_id[duplicated(participants_oxf$child_id)]) |>
        pivot_wider(
            id_cols = child_id,
            names_from = session_n,
            values_from = age,
            names_repair = janitor::make_clean_names
        ) |>
        rowwise() |>
        mutate(diff_min = x2 - x1) |>
        ungroup() |>
        summarise(diff_min = min(diff_min, na.rm = TRUE))
})

@@

```{r tbl-participants}
#| label: tbl-participants
#| tbl-cap: "Demographic and linguistic profile of testing sessions. The number of excluded testing sessions and participants is indicated between parentheses."
participants |>
    unnest_wider(doe) |>
    select(-matches("others")) |>
    arrange(lp, test_language) |>
    distinct(session_id, child_id, .keep_all = TRUE) |>
    mutate(
        study = if_else(location == "Oxford",
            "Study 1",
            "Study 2"
        ),
        test_language = paste0(test_language, " dominant"),
        lp = gsub(" \\(English\\)", "", lp)
    ) |>
    inner_join(attrition_participants,
        by = join_by(session_id)
    ) |>
    summarise(
        n = n(),
        n_participants = length(unique(child_id)),
        across(age, lst(mean, sd)),
        across(matches("doe_"), lst(mean, sd)),
        # age_vctr = list(age),
        .by = c(lp, study, test_language, is_valid_participant)
    ) |>
    pivot_wider(
        id_cols = c(lp, study, test_language),
        names_from = is_valid_participant,
        values_from = is.numeric
    ) |>
    select(
        lp, study, test_language, starts_with("n_"), matches("TRUE"),
        n_FALSE, n_participants_FALSE
    ) |>
    rename_with(\(x) gsub("_TRUE", "", x)) |>
    gt(
        rowname_col = "test_language",
        groupname_col = c("study", "lp")
    ) |>
    grand_summary_rows(
        columns = matches("age"),
        fns = list(Total ~ mean(.)),
        fmt = ~ fmt_number(., decimals = 2)
    ) |>
    grand_summary_rows(
        columns = c(n, n_participants),
        fns = list(Total ~ sum(.)),
        fmt = ~ fmt_integer(.)
    ) |>
    # gt_plt_dist(age_vctr, type = "histogram") |>
    fmt_number(is.numeric, decimals = 2) |>
    fmt_integer(c(n, n_participants, n_FALSE, n_participants_FALSE)) |>
    fmt_number(matches("doe"), decimals = 2, scale_by = 100) |>
    cols_merge(c(n, n_FALSE), pattern = "{1} ({2})") |>
    cols_merge(c(n_participants, n_participants_FALSE),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(age_mean, age_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(doe_catalan_mean, doe_catalan_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(doe_english_mean, doe_english_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(doe_spanish_mean, doe_spanish_sd),
        pattern = "{1} ({2})"
    ) |>
    tab_spanner("Sample size", starts_with("n")) |>
    tab_spanner("Age (months)", matches("age")) |>
    tab_spanner("Catalan", matches("catalan")) |>
    tab_spanner("English", matches("english")) |>
    tab_spanner("Spanish", matches("spanish")) |>
    tab_spanner("Degree of Exposure (%)", matches("doe")) |>
    cols_label(
        n = "Test sessions",
        n_participants = "Participants",
        age_mean = "M (SD)",
        doe_catalan_mean = "M (SD)",
        doe_english_mean = "M (SD)",
        doe_spanish_mean = "M (SD)"
    ) |>
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels()
    ) |>
    tab_style(
        cell_text(style = "italic"),
        cells_column_spanners()
    )
```

```{r vocab-values-oxf}
vocabulary_oxf <- inner_join(vocabulary,
    select(participants_oxf, session_id),
    by = join_by(session_id)
)

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles."
#| fig-width: 5
#| fig-height: 4
#| out-width: 70%
vocab_plot_data <- vocabulary |>
    inner_join(filter(attrition_participants, is_valid_participant),
        by = join_by(session_id)
    ) |>
    inner_join(
        select(
            participants, location, lp, age,
            child_id, session_id, age_group
        ),
        by = join_by(child_id, session_id)
    ) |>
    filter(is_valid_participant) |>
    select(
        child_id, session_id, vocab_id, is_imputed, l1_count,
        lp, location, age_group, age
    ) |>
    mutate(
        location = factor(location,
            levels = c("Oxford", "Barcelona")
        ),
        lp = gsub(" \\(English\\)", "", lp),
        side = case_when(
            location == "Oxford" ~ "both",
            lp == "Monolingual" ~ "left",
            lp == "Bilingual" ~ "right"
        ),
        study = if_else(location == "Oxford",
            "Study 1 (English)", "Study 2 (Catalan/Spanish)"
        )
    )

vocab_plot_means <- vocab_plot_data |>
    summarise(
        l1_count_sd = sd(l1_count, na.rm = TRUE),
        l1_count = mean(l1_count, na.rm = TRUE),
        n = n(),
        .by = c(age, side, lp, study)
    ) |>
    mutate(
        .lower = l1_count - (l1_count_sd / sqrt(n)),
        .upper = l1_count + (l1_count_sd / sqrt(n))
    )

vocab_plot_data |>
    filter(study == "Study 1 (English)") |>
    ggplot(aes(age, l1_count,
        side = side
    )) +
    geom_point(
        size = 2,
        shape = 1,
        colour = "grey",
        alpha = 3 / 4,
        stroke = 1
    ) +
    geom_smooth(
        method = "lm",
        formula = "y ~ x",
        colour = "black",
        fill = "black"
    ) +
    labs(
        x = "Age (months)",
        y = "Vocabulary size",
        colour = "Group",
        fill = "Group"
    ) +
    scale_y_continuous(
        limits = c(0, 575),
        breaks = seq(0, 1e3, 100)
    ) +
    scale_x_continuous(breaks = seq(0, 50, 2)) +
    theme(
        panel.grid.major.y = element_line(
            linetype = "dotted",
            colour = "grey"
        ),
        legend.position = "top",
        legend.title = element_blank()
    )
```

@@

Participants were presented with 32 trials in random order, which belonged to two conditions: *Related* and *Unrelated* trials. In *Related* trials (*n* = 16), the English label of the prime was phonologically related to the target label, sharing phonological onset (e.g., /**\textipa{"t}**\textipa{\*ri:}/$_{\text{ENG}} [\text{tree}]$--/**\textipa{"t}**\textipa{\*r2k}/$_{\text{ENG}} [\text{truck}]$). In *Unrelated* trials, the prime and target labels did not share phonological onset (e.g., /\textipa{"dO:}/$_{\text{ENG}} [\text{door}]$--/\textipa{"s6k}/$_{\text{ENG}} [\text{sock}]$). Especial attention was paid to avoiding semantic or taxonomic relationships between prime and target words, and between prime and distractor words. Distractors were always phonologically unrelated to the prime and target labels in the same trial.

@@

```{r stim-stats}
durations <- stimuli |>
    summarise(
        across(
            duration,
            lst(mean, sd, min, max)
        ),
        .by = c(test_language)
    ) %>%
    mutate(across(is.numeric, \(x) format(round(x * 1e3, 2), big.mark = ","))) %>%
    split(janitor::make_clean_names(.$test_language))
```

@@

```{r tbl-stimuli}
#| label: tbl-stimuli
#| tbl-cap: "Summary of stimuli properties by trial type. Values are summarised using the mean and the standard deviation (between parentheses)."
stimuli |>
    unnest_wider(c(freq, nphon, familiarity)) |>
    summarise(
        across(
            c(matches("n_phon|freq_|nphon|familiarity")),
            list(
                mean = \(x) mean(x, na.rm = TRUE),
                sd = \(x) sd(x, na.rm = TRUE)
            )
        ),
        .by = c(trial_type, test_language)
    ) |>
    select(-matches("distractor")) |>
    mutate(test_language = if_else(test_language == "English",
        paste0("Study 1: ", test_language),
        paste0("Study 2: ", test_language)
    )) |>
    arrange(test_language) |>
    gt(
        rowname_col = "trial_type",
        groupname_col = "test_language"
    ) |>
    # summary_rows(groups = c("Unrelated", "Non-cognate", "Cognate"),
    # 			 columns = is.numeric,
    # 			 fns = list(Total ~ mean(.)),
    # 			 fmt = ~ fmt_number(., decimals = 2)) |>
    grand_summary_rows(
        columns = matches("phon|freq"),
        fns = list(Total ~ mean(.)),
        fmt = ~ fmt_number(., decimals = 2)
    ) |>
    grand_summary_rows(
        columns = matches("familiarity"),
        fns = list(Total ~ mean(.)),
        fmt = ~ fmt_number(., decimals = 2, scale_by = 100)
    ) |>
    tab_spanner("# Phonemes", matches("nphon")) |>
    tab_spanner("Frequency (Zipf)", matches("freq")) |>
    tab_spanner("Familiarity (%)", matches("familiarity")) |>
    fmt_number(is.numeric, decimals = 2) |>
    fmt_number(matches("familiarity"), decimals = 2, scale_by = 100) |>
    cols_merge(c(freq_prime_mean, freq_prime_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(freq_target_mean, freq_target_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(familiarity_prime_mean, familiarity_prime_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(familiarity_target_mean, familiarity_target_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(nphon_prime_mean, nphon_prime_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(nphon_target_mean, nphon_target_sd),
        pattern = "{1} ({2})"
    ) |>
    cols_label(
        trial_type = "",
        familiarity_prime_mean = "Prime",
        familiarity_target_mean = "Target",
        freq_prime_mean = "Prime",
        freq_target_mean = "Target",
        nphon_prime_mean = "Prime",
        nphon_target_mean = "Target"
    ) |>
    tab_style(
        list(cell_text(style = "italic")),
        list(
            cells_column_labels(),
            cells_column_spanners(),
            cells_grand_summary()
        )
    )
```

@@

We defined a time window of interest from 200 ms after target and distractor pictures onset until the end of the trial at 2,000 ms when both pictures disappeared from screen. The first 200 ms of the test phase were discarded to avoid modelling fixations driven by processes other than auditory word recognition [@fernald1998rapid; @fernald2001half]. Missing eye-tracker samples were interpolated using the last-observation-carried-forward [see @zettersten2022peekbank for a similar approach], with a maximum of 20 maximum consecutive missing samples being interpolated (an equivalent of `r round(20*1000/120, 2)`). Target looking probability was calculated as the empirical logit, using the number of samples inside the time bin in which the participant was looking at the target and distractor AOI (see @eq-elogit) [@barr2008analyzing; @agresti2012categorical], as follows:

@@

After trials that matched any of the aforementioned exclusion criteria from the data set, we excluded participants who did not provide at least two valid trials in each condition (*n* = `r n_exc_participants`), and participants with a vocabulary size lower than 42, which corresponds to 10% of the words in the OCDI vocabulary checklist (*n* = `r n_exc_participants_vocab`). The final data set included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final data set, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), and `r n_trials_condition$non_cognate_2$n + n_trials_condition$cognate_2$n` were *Related*  trials(`r n_trials_condition$non_cognate$n + n_trials_condition$cognate$n` previously excluded).

@@

n_lp <- attrition |>
    distinct(is_valid_participant, lp, session_id) |>
    count(is_valid_participant, lp) |>
    pivot_wider(
        names_from = is_valid_participant,
        values_from = n,
        names_repair = janitor::make_clean_names
    ) |>
    mutate(
        n_participants_true = true,
        n_participants_false = false
    ) |>
    select(lp, true, false)

attrition |>
    mutate(study = if_else(location == "Oxford",
        "Study 1: English (Oxford)",
        "Study 2: Catalan/Spanish (Barcelona)"
    )) |>
    summarise(
        n_trials_true = sum(is_valid_trial),
        n_trials_false = sum(!is_valid_trial),
        .by = c(lp, trial_type, study)
    ) |>
    arrange(study, desc(lp)) |>
    pivot_wider(
        id_cols = c(study, lp),
        names_from = c(trial_type),
        values_from = matches("trials"),
        names_repair = janitor::make_clean_names
    ) |>
    left_join(n_lp, by = join_by(lp)) |>
    select(study, lp, true, false, matches("n_trials")) |>
    gt(
        rowname_col = "lp",
        groupname_col = "study"
    ) |>
    tab_spanner("Participants", matches("participant")) |>
    tab_spanner("Trials", matches("trial")) |>
    fmt_integer(everything()) |>
    cols_merge(c(true, false), pattern = "{1} ({2})") |>
    cols_merge(c(n_trials_true_cognate, n_trials_false_cognate),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(n_trials_true_non_cognate, n_trials_false_non_cognate),
        pattern = "{1} ({2})"
    ) |>
    cols_merge(c(n_trials_true_unrelated, n_trials_false_unrelated),
        pattern = "{1} ({2})"
    ) |>
    cols_label(
        true = "Participants",
        n_trials_true_cognate = "Cognate",
        n_trials_true_non_cognate = "Non-cognate",
        n_trials_true_unrelated = "Unrelated"
    ) |>
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels()
    ) |>
    tab_style(
        cell_text(align = "left"),
        cells_stub()
    ) |>
    grand_summary_rows(
        fns = list(label = md("*N*")) ~ sum(.),
        fmt = ~ fmt_integer(.)
    )
```

@@

r_version <- glue::glue(r_info$R.version$major,
    r_info$R.version$minor,
    .sep = "."
)
stan_version <- "2.33.0"

@@

We used Bayesian Hierarchical General Additive Models (HGAMs) to model the data [@pedersen2019hierarchical], using a Gaussian distribution to model the the logit of target looking. First, we fit a model ($\mathcal{M}_0$) that included the main effects of *Condition* and *Age* as fixed effects in the model. We set an *a priori* contrast for the *Condition* predictor [@schad2020capitalize], comparing *Unrelated* and *Related* trials (sum-coded as `-0.5` and `+0.5`. Before entering the model, the *Age* predictor was standardised by subtracting from each observation the mean of the predictor, and dividing the result from the standard deviation of the predictor. We included the variable *Session*---which indexes individual testing sessions that may belong to the same participant---as grouping variable, nested within the *Participant* grouping variable---which indexes a distinct participant. This nested random effects structure incorporates the longitudinal design of data collection, in which multiple participants were tested more than once at different ages. We added by-session intercepts and *Condition* slopes, and by-participant intercepts and *Age* slopes. To model the time course of target looking across time bins, we included B-splines for the main effect of *Time*, and for the *Condition* predictor [@wood2017generalized]. For both splines, we specified $k = 8$ basis functions or *knots*. @eq-model shows a formal implementation of the model.

@@

```{r coefs-oxf}
var_labels <- c(
    "b_Intercept" = "Intercept",
    "b_age_std" = "Age (+1 SD)",
    "b_condition1" = "Condition (U vs. R/NC)",
    "b_condition2" = "Condition (R/NC vs. R/C)"
)

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_0,
    data_bcn,
    vars_dict = var_labels,
    rope_interval = rope_interval
)

coefs <- model_summary |>
    mutate(
        .value = sprintf("%.3f", .value),
        .ci = sprintf("[%.3f, %.3f]", .lower, .upper),
        .rope = scales::percent(.rope, accuracy = 0.01)
    ) |>
    split(janitor::make_clean_names(model_summary$.variable))
```
